Hierarchical Dispatching and Overriding

### 2017-10-03 NewVersion/sections/Auxiliary.tex

\section{Auxiliary Definitions}\label{subsec:auxdefs}
In this section, we present the auxiliary definitions used in our formalization in 
Figure~\ref{fig:auxfunc}. To begin with, we
introduce the basic functions: $\ext, \overrideSet$ and $\prune$. $\ext(I, J)$
simply indicates that interface $I$ directly extends interface $J$. 
$\overrideSet(I, J)$ returns a set of methods defined in $I$ that have ``$\kwoverride \; J$''
in their signatures. Notice that $\overrideSet(I, I)$ is a special representative of
the ``originally-defined'' method set from $I$. The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. Finally in the returned set,
none of them has a subtyping to one another, since all super types have been removed.

@@

    \inferrule* [left=]
    { originalIntfs = \; \{K \mid \subt{I}{K}, \; \subt{K}{J} \; \lor \; \subt{J}{K}, \; $and$\; m \in \overrideSet(K, K) \} }
    {\mostSpecific(m, I, J) = \prune(originalIntfs) }

    \inferrule* [left=]
        { overrideIntfs = \; \{ K\mid I <: K, \; K <: J\; $and$ \;
            m \in \overrideSet(K, J)  \} }
        {\mostSpecificOverride(m, I, J) = \prune(overrideIntfs)}

@@

    \inferrule* [left=]
    {   \kwinterface \; I \; \kwextends \; \overline{I} \; \{ I_e \; m(\overline{I_x} \; \overline{x}) \;
        \kwoverride \; J \ldots \} }
    {m \in \overrideSet(I, J)}

>>>>

\section{Auxiliary Definitions}\label{subsec:auxdefs}
In this section, we present the auxiliary definitions used in our formalization in 
Figure~\ref{fig:auxfunc}. To begin with, we
introduce the basic functions: $\ext, I.m\,\kwoverride\,J$ and $\prune$. $\ext(I, J)$
simply indicates that interface $I$ directly extends interface $J$. 
$\overrideSet(I, J)$ returns a set of methods defined in $I$ that have ``$\kwoverride \; J$''
in their signatures. \marco{text need to change since notation is changed.} Notice that $\overrideSet(I, I)$ is a special representative of
the ``originally-defined'' method set from $I$. The $\prune$ function takes a set of
types, and filters out those that have subtypes in the same set. Finally in the returned set,
none of them has a subtyping to one another, since all super types have been removed.

@@

    \inferrule* [left=]
    { origins = \; \{K \mid \subt{I}{K}, \; \subt{K}{J} \; \lor \; \subt{J}{K}, \; $and$\; K.m\,\kwoverride\,K \} }
    {\mostSpecific(m, I, J) = \prune(origins) }

    \inferrule* [left=]
        { overrides = \; \{ K\mid I <: K, \; K <: J\; $and$ \;
            K.m\,\kwoverride\,J  \} }
        {\mostSpecificOverride(m, I, J) = \prune(overrides)}

@@

    \inferrule* [left=]
    {   \kwinterface \; I \; \kwextends \; \overline{I} \; \{ I_e \; m(\overline{I_x} \; \overline{x}) \;
        \kwoverride \; J \ldots \} }
    {I.m\,\kwoverride\,J}

### 2017-10-03 NewVersion/sections/Overview.tex

\subsection{Problem 1: Unintentional Method Conflicts}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| in \lstinline|Deck| simply
prints out a message.

@@

interface Drawable {
  JFrame draw() { // draws something on the screen
    JFrame frame = new JFrame();
    frame.setVisible(true);
    return frame;
  }
}
\end{lstlisting}\vspace{3pt}
Note that the two \lstinline|draw| methods have the same names and parameter types,
but the return types can be different. In \lstinline|Deck|, \lstinline|println| is a
library function, loaded by the prelude for example. Since the only operation is messaging,
the method has \lstinline|void| return type (we could define an interface called \lstinline|Void|
and return an object of it instead; in many languages it is simply a primitive type). On the other hand, for \lstinline|Drawable|
the program has to create a blank canvas by \lstinline|JFrame|, again a pre-defined type:

\vspace{3pt}\begin{lstlisting}
interface JFrame {
  void setVisible(boolean b) {...}
  ...
}
\end{lstlisting}\vspace{3pt}
Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

\vspace{3pt}\begin{lstlisting}
interface DrawableDeck extends Drawable, Deck {...} 
\end{lstlisting}\vspace{3pt}
The point of using multiple inheritance is surely for composing the features from various 
components and to achieve code reuse, as supported by many mainstream OO
languages. Nevertheless at this point, languages like Java simply treat the two \lstinline|draw| methods
as the same, and hence throw a compile error on \lstinline|DrawableDeck| about the conflict, though accidentally.

Now one may quickly come up with a workaround, which is creating a new \lstinline|draw| method in \lstinline|DrawableDeck| to
explicitly override the old ones. However, the two conflicting methods have totally different functionalities, merging
them into one does not make any sense. This non-solution will hide the
old methods and break independent extensibility.

\subsection{Potential Fixes}

There are several other workarounds
that come to our mind. We briefly discuss those potential fixes and
workarounds next:
\begin{itemize}
  \item \textbf{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields (or field methods) with
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. This avoids method conflicts. Nevertheless, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.
  \item \textbf{II. Refactor Drawable and/or Deck to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However this approach is non-modular, as it requires 
  modifying existing code, and may not be possible if code is unavailable.
  \item \textbf{III. Method exclusion/renaming in traits.} Some trait models
  support method exclusion/renaming. Those features
   can eliminate conflicts, although most
  programming languages do not support them. In a traditional OO system,
  they can break the subtyping relationship. Moreover, in
  contrast with exclusion, renaming can indeed preserve both conflicting
  behaviours, however, it is cumbersome in practice, as introducing new
  names can affect other code blocks.
  \item \textbf{IV. Static dispatch.} Some other languages like C++ have
  direct support for the triangle inheritance. For instance, C++ allows both
  conflicting methods to coexist, and resolves the ambiguity by static dispatch:
  \vspace{3pt}\begin{lstlisting}[language=c++]
  class Deck { public: void draw() {...} };
  class Drawable { public: JFrame draw() {...} };
  class DrawableDeck : public Drawable, public Deck {};

@@

So far it seems that static dispatch is the simplest and most straightforward approach,
as long as the language model accepts unintentionally conflicting methods. However, dynamic
dipatching appears to be more common in OO programming in order for code reuse. Following the previous example,
we reimplement \lstinline|Deck| with more features:

@@

\vspace{3pt}\begin{lstlisting}
interface SafeDeck extends Deck {
  boolean isEmpty() {...}
  void draw() { // overriding
    if (isEmpty()) println("The deck is empty.");
    else println("Draw a card");
  }
}
\end{lstlisting}\vspace{3pt}
With static dispatch, it is tedious but we have to copy the \lstinline|shuffleAndDraw| code into \lstinline|SafeDeck|,
for the adaption to the new \lstinline|draw|. Yet dynamic dispatching immediately saves us from the duplicate work,
since the method becomes automatically adaptive. Nevertheless, as seen before, dynamic dispatch would take us back to problem
1, for instance, when we
have the UML in Figure~\ref{fig:drawableloggingdeck} and the following code:

>>>>

\subsection{Problem 1: Unintentional Method Conflicts}
Suppose that two components \lstinline|Deck| and \lstinline|Drawable| 
have been developed in a system. \lstinline|Deck| represents a deck
of cards and defines a method \lstinline|draw| for drawing a card from the
deck.  \lstinline|Drawable| is an interface for graphics that
can be drawn and also includes a method called \lstinline|draw| for
visual display. For simple illustration, the default implementation of
\lstinline|draw| in \lstinline|Drawable| only creates a blank canvas
on the screen, while the \lstinline|draw| method in \lstinline|Deck| simply
prints out a message.

@@

interface Drawable {
  JFrame draw() { // draws something on the screen
    JFrame frame = new JFrame();
    frame.setVisible(true);
    return frame;
  }
}
\end{lstlisting}\vspace{3pt}
Note that the two \lstinline|draw| methods have the same names and parameter types,
but the return types can be different. In \lstinline|Deck|,
\lstinline|draw| uses \lstinline|println|, which is a
library function. 
The method returns \lstinline|void|. Note that \lstinline|void| is
unsupported in our formalization, but we could have also defined an interface called \lstinline|Void|
and return an object of that type instead. For the method \lstinline|draw| in
\lstinline|Drawable|
the program has to create a blank canvas by \lstinline|JFrame|, which
is a pre-defined type.
\begin{comment}
\vspace{3pt}\begin{lstlisting}
interface JFrame {
  void setVisible(boolean b) {...}
  ...
}
\end{lstlisting}\vspace{3pt}
\end{comment}
Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

\vspace{3pt}\begin{lstlisting}
interface DrawableDeck extends Drawable, Deck {...} 
\end{lstlisting}\vspace{3pt}
The point of using multiple inheritance is for composing the features from various 
components and to achieve code reuse, as supported by many mainstream OO
languages. Nevertheless at this point, languages like Java simply treat the two \lstinline|draw| methods
as the same, and hence they throw a compile error on \lstinline|DrawableDeck| about the conflict, though accidentally.

Now one may quickly come up with a workaround, which is to create a new \lstinline|draw| method in \lstinline|DrawableDeck| to
explicitly override the old ones. However, the two conflicting methods
have totally different functionalities. Therefore merging
them into one does not make any sense. This non-solution will hide the
old methods and break independent extensibility.

\paragraph{{\bf Potential Fixes:}}
There are several other workarounds
that come to our mind. We briefly discuss those potential fixes and
workarounds next:
\begin{itemize}
  \item \textbf{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields (or field methods) with
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. This avoids method conflicts. Nevertheless, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.
  \item \textbf{II. Refactor Drawable and/or Deck to rename the methods.} If
  the source code for \lstinline|Drawable| or \lstinline|Deck| is available
  then it may be possible to rename one of the \lstinline|draw|
  methods. However this approach is non-modular, as it requires 
  modifying existing code, and may not be possible if code is unavailable.
  \item \textbf{III. Method exclusion/renaming in traits.} Some trait models
  support method exclusion/renaming. Those features
   can eliminate conflicts, although most
  programming languages do not support them. In a traditional OO system,
  they can break the subtyping relationship. Moreover, in
  contrast with exclusion, renaming can indeed preserve both conflicting
  behaviours, however, it is cumbersome in practice, as introducing new
  names can affect other code blocks.
  \item \textbf{IV. Static dispatch.} Some other languages like C++ have
  direct support for the triangle inheritance. For instance, C++ allows both
  conflicting methods to coexist, and resolves the ambiguity by static dispatch:
  \vspace{3pt}\begin{lstlisting}[language=c++]
  class Deck { public: void draw() {...} };
  class Drawable { public: JFrame draw() {...} };
  class DrawableDeck : public Drawable, public Deck {};

@@

\bruno{Why have we eliminated problem 2 and included it in Problem 1?
Also why are we not showing how to actually solve it with our approach
now? I what we had previously made more sense. Please finish up this
section; show our solution, mention that is is quite similar (and
inspired by C++), but finish it saying that there are other problems
that C++ does not address.}
\subsection{Problem 2: Dynamic Dispatching}
So far it seems that static dispatch is the simplest and most straightforward approach,
as long as the language model accepts unintentionally conflicting methods. However, dynamic
dispatching appears to be more common in OO programming in order for code reuse. Following the previous example,
we reimplement \lstinline|Deck| with more features:

@@

\vspace{3pt}\begin{lstlisting}
interface SafeDeck extends Deck {
  boolean isEmpty() {...}
  void draw() { // overriding
    if (isEmpty()) println("The deck is empty.");
    else println("Draw a card");
  }
}
\end{lstlisting}\vspace{3pt}
With static dispatch, we have to copy the \lstinline|shuffleAndDraw| code into \lstinline|SafeDeck|,
for the adaptation to the new \lstinline|draw|. Yet dynamic dispatching immediately saves us from the duplicate work,
since the method becomes automatically adaptive. Nevertheless, as seen before, dynamic dispatch would take us back to problem
1, for instance, when we
have the UML in Figure~\ref{fig:drawableloggingdeck} and the following code:
